\textcolor{violet}{Reverse Correlation Uncovers More Complete Tinnitus Spectra \vspace{0.25cm}}

### 2023-01-05 paper/short/main_wc995.tex
Cleanup

\section{Introduction}
% INTRO PARAGRAPH 1/3
\IEEEPARstart{T}{INNITUS}\textemdash{}the perception of sound
in the absence of any corresponding external stimulus\textemdash{}affects up to 50 million people in the U.S. \cite{bhattPrevalenceSeverityExposures2016,nationalinstituteondeafnessandothercommunicationdisordersQuickStatisticsHearing},
a third of whom experience functional cognitive impairment and substantial reduction in quality of life
\cite{ahmedImpactTinnitusPerception2017,nondahlImpactTinnitusQuality2007,tegg-quinnImpactTinnitusCognition2016}.
Clinical guidelines for tinnitus management frequently involve targeted exposure to tinnitus-like external sounds, either as a form of habituation therapy (termed \textit{sound therapy}), or in the context of Cognitive Behavioral Therapy \cite{tunkelClinicalPracticeGuideline2014}.
Critically, sound therapy is shown to improve with knowledge of the patient's internal tinnitus experience \cite{davisNeuromonicsTinnitusTreatment2007,landgrebeMethodologicalAspectsClinical2012,nickelOutcomeResearchMusic2005,okamotoListeningTailormadeNotched2010,schaetteAcousticStimulationTreatments2010,steinInhibitioninducedPlasticityTinnitus2015,tassCounteractingTinnitusAcoustic2012,wangStateArtSound2020}.
However, existing strategies for characterizing tinnitus sounds, such as Pitch Matching (PM) methods, are best suited for patients who experience sounds resembling pure tones (\eg ringing).
This group may represent only between 20-50\% of all tinnitus patients \cite{turnerjrAuditoryDysfunctionTinnitus1990,henryAuditoryTestResult2009,ukaegbeTinnitusItsEffect2017}, with others experiencing non-tonal (\eg buzzing, roaring) tinnitus sounds.
There is a pressing need for more robust tinnitus characterization methods \cite{henryTinnitusEpidemiologicPerspective2020,henryMeasurementTinnitus2016,norenaPsychoacousticCharacterizationTinnitus2002,robertsResidualInhibitionFunctions2006}.

% INTRO PARAGRAPH 2/3
Tinnitus sounds are presumed to be both complex and heterogeneous
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,henryTinnitusEpidemiologicPerspective2020,meikleTinnitusArchiveArchive2004},
although little has been firmly established about the specific characteristics.
Therefore, we base our approach on \textit{reverse correlation} (RC),
an established behavioral method
\cite{ahumadaStimulusFeaturesSignal1971,gosselinSuperstitiousPerceptionsReveal2003,brimijoinInternalRepresentationVowel2013}
for probing internal representations of visual and auditory percepts that is unconstrained by prior knowledge about the percepts themselves.
RC asks subjects to render subjective judgments (\eg “does this sound like your tinnitus?”)
over random stimuli, 
from which unbiased estimates of the percepts can be produced.
RC is closely related to Wiener theory \cite{wienerNonlinearProblemsRandom1966},
which has also inspired ``white-noise'' approaches to system characterization in physiology
\cite{ringachReverseCorrelationNeurophysiology2004,marmarelisWhiteNoiseMethodSystem1978} and engineering more broadly
\cite{ljungMeasureLackFit1978}.

% INTRO PARAGRAPH 3/3
Here, we validate RC as a method for characterizing all component
frequencies of non-tonal tinnitus sounds
by estimating the full psychoacoustic tinnitus spectrum (PTS).
To that end, we asked normal-hearing participants to complete an augmented RC experiment, comparing random stimuli to a target tinnitus-like sound.
Target sounds were subsequently used as the basis for validating the quality of the estimated PTS.
Our results demonstrate, for the first time, that tinnitus-like sounds with complex spectra can be accurately estimated using RC. 
This discovery may open the door to developing a clinical assay for characterizing the full range of sounds experienced by tinnitus patients.

@@

\section{Materials and Methods }
% To use this template, type over sections of the template or cut and paste from another document. Highlight a section that you want to designate with a certain style, then select the appropriate name on the style menu. The style will adjust your fonts and line spacing. Do not change the font sizes or line spacing to squeeze more text into a limited number of pages. Use italics for emphasis; do not underline. 
% IEEE will do the final formatting of your paper. However, you should format your manuscript for submission to IEEE OJEMB for review and also for preparing for preprint that will be e-published several days after you upload the final version of an accepted manuscript.

Software code used for the experiments and analysis was written in MATLAB and is freely available at \protect\url{https://github.com/alec-hoyland/tinnitus-reconstruction/}.

@@

Stimuli were composed of $b=8$ Mel-spaced frequency bins,
which divide the frequency spectrum between $[100,~13,000]$
Hz into continuous segments of equal amplitude (\ie ``rectangular'' bins).
Reconstruction detail increases with $b$, but $b=8$ bins provides a good approximation to the chosen target sounds (see Fig. \ref{fig:spectra}).

For each stimulus, $[2,7]$ bins were randomly ``filled'' with power $0$ dB. ``Unfilled'' bins were assigned a power of $-100$ dB. 
Assigning random phase, the inverse Fourier transform of the spectrum yields a $500$-ms stimulus waveform.

@@

Two tinnitus-like target sounds (``buzzing'' and ``roaring'') from online examples maintained by the American Tinnitus Association
\cite{MeasuringTinnitusAmerican2022} were downloaded and truncated to $500$-ms in duration. These sounds were selected for the complexity and complementary nature of their frequency spectra.

@@

Trials followed an $A-X$ paradigm (Fig. \ref{fig:experimentdiagram}), where $n=10$ subjects with self-reported normal hearing listened to the same target sound ($A$) followed by a random stimulus ($X$) for each trial.
The task is designed to mimic the process by which a tinnitus patient might compare a random stimulus to their own tinnitus percept.

Subjects were told that some stimuli had the target sound embedded in them, and were instructed to respond ``yes'' to any such stimuli, and ``no'' otherwise.
Subjects performed trials in blocks of $100$, with breaks between blocks.
Subjects completed two ($2$) blocks of trials per given target, $A$.
The $400$ total trials took subjects 10-20 minutes to complete.

@@

where $x \in \mathbb{R}^\mathrm{b}$ represents the subject's internal representation of interest (\eg their tinnitus percept).
This model can be inverted, yielding the following formula:

@@

The $A-X$ paradigm allows for direct validation of the reconstructions $\hat{x}_{buzzing}$ and $\hat{x}_{roaring}$ against their corresponding target sounds.
To that end, we represent the power spectra of the target sounds as vectors $s_{buzzing} \in \mathbb{R}^\mathrm{b}$ and $s_{roaring} \in \mathbb{R}^\mathrm{b}$ using the same $b=8$ frequency bins used in stimulus generation.
The power in each bin is set as the mean power level of frequencies within that bin.
Pearson's $r$ between $s_{buzzing}$ and $s_{roaring}$ and their corresponding reconstructions ($\hat{x}_{buzzing}$ and $\hat{x}_{roaring}$) represents reconstruction accuracy.
One-sample t-tests were performed on the mean Fisher-transformed $r$ values across subjects to assess significant differences from zero.

\subsection{Synthetic and Random Subjects}

To establish bounds on human performance,
additional experiments were run \textit{in-silico} with two simulated subjects that
represent participants who always give either \textit{ideal} or \textit{random} responses.
Each experiment ran for $n=200$ trials and was repeated $1000$ times.

@@

% NB NOTE: What is s??
for $i \in 1, ..., n$, where $Q(x, y)$ is the quantile function for $x \in [0, 1]$ of the similarity calculation $\Psi s$, and
$\Psi_i$ is the $i^\mathrm{th}$ column of $\Psi$.
The ideal subject represents one who has precise knowledge of every stimulus and where the reconstruction algorithm mirrors \ref{eq:linreg}.

The \textit{random} subject chooses responses at random:

\begin{equation}
    r_i =
    \begin{cases}
        1 & \text{if  } X > 0.5 \\
        -1 & \text{otherwise}
    \end{cases}
\end{equation}

where $X \in [0, 1]$ is a uniform random variable.

@@

Figure \ref{fig:reconstructions} shows the distribution of Pearson's $r$ (\ie reconstruction accuracy) from human, ideal, and random subject responses in the buzzing and roaring experimental conditions.
Accuracies from human responses were generally higher than those from random subject responses,
with some accuracies approaching those from ideal subject responses.
Mean accuracy from the random subject responses was $0$ in all conditions,
while mean accuracy from human responses was significantly different from $0$ in all conditions: buzzing: $t(9) = 5.766$, $p < 0.001$; roaring: $t(9) = 5.76$, $p < 0.001$; combined: $t(19) = 7.542$, $p < 0.001$.
The buzzing results vary more than the roaring results. The buzzing ground truth spectrum contains significant high-frequency content,
suggesting that the task may be more difficult for subjects with minor high-frequency hearing loss.
Figure \ref{fig:spectra} plots example reconstructions from human responses with the best reconstruction accuracies overlaid on the target sound spectra,
showing that those reconstructions capture many of the salient features of the target.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{reconstruction_violin_1.eps}
    \caption{Reconstruction accuracy for human subjects is significantly above baseline,
    but is not optimal. Reconstruction accuracies of random, human, and ideal subjects
    are shown as violin plots with box plots overlaid.
    The median is a white dot, the ordinate of the blue points are the Pearson's $r$ values.}
    \label{fig:reconstructions}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{reconstruction_spectra.eps}
    \caption{Reconstructions of the PTS capture many salient features of tinnitus spectra.
    On each set of axes, the black trace indicates the target,
    while colored traces plot exemplar reconstructions from human subjects' data.
    In the top row, the target and reconstructions are shown as
    standardized power levels within each frequency bin.
    In the bottom row, those standardized power levels are mapped from the $8$-dimensional bin domain
    to a $11025$-dimensional frequency domain.
    The unbinned power spectral density of the targets is also shown in gray.}
    \label{fig:spectra}
\end{figure}

@@

% Promise
Our results show that RC can accurately reconstruct the frequency spectrum of complex, non-tonal tinnitus-like sounds.
This indicates that RC may also be useful for developing a clinical assay to characterize a wider variety of tinnitus percepts than currently possible.
RC is purely behavioral, requires no specialized equipment, and has marginal setup.
Moreover, subjects in the present study were able to complete the required number of trials within 20 minutes,
indicating that this procedure could be conducted in a single clinical visit.
Future work will focus on validating this approach in patients suffering from tinnitus.

% Limitations & Future Work
Reconstruction accuracies observed here are well below those provided by an ideal subject's responses,
which may be attributed to noisy human response characteristics universally observed in applications of RC.
This may be mitigated by optimizing the experimental protocol, stimulus generation, and reconstruction method.
For example, recent approaches to improving RC reconstruction methods have been shown to boost efficiency, noise robustness and overall accuracy \cite{roopCompressiveSensingApproach2021,comptonStimulusWhiteningImproves2022a},
and may also be applicable in the context of tinnitus.

>>>>

\section{Introduction}
% INTRO PARAGRAPH 1/3
\IEEEPARstart{T}{INNITUS}\textemdash{}the perception of sound
in the absence of any corresponding external stimulus\textemdash{}affects up to 50 million people in the U.S. \cite{bhattPrevalenceSeverityExposures2016,nationalinstituteondeafnessandothercommunicationdisordersQuickStatisticsHearing},
a third of whom experience functional cognitive impairment and substantial reduction in quality of life
\cite{ahmedImpactTinnitusPerception2017,nondahlImpactTinnitusQuality2007,tegg-quinnImpactTinnitusCognition2016}.
Clinical guidelines for tinnitus management frequently involve targeted exposure to tinnitus-like sounds, either as a form of habituation therapy (termed \textit{sound therapy}), or using Cognitive Behavioral Therapy \cite{tunkelClinicalPracticeGuideline2014}.
Critically, sound therapy is shown to improve with knowledge of the patient's internal tinnitus experience \cite{davisNeuromonicsTinnitusTreatment2007,landgrebeMethodologicalAspectsClinical2012,nickelOutcomeResearchMusic2005,okamotoListeningTailormadeNotched2010,schaetteAcousticStimulationTreatments2010,steinInhibitioninducedPlasticityTinnitus2015,tassCounteractingTinnitusAcoustic2012,wangStateArtSound2020}.
However, existing strategies for characterizing tinnitus sounds, such as Pitch Matching (PM) methods, are best suited for patients whose tinnitus resembles pure tones (\eg ringing), which may represent only $20-50\%$ of patients \cite{turnerjrAuditoryDysfunctionTinnitus1990,henryAuditoryTestResult2009,ukaegbeTinnitusItsEffect2017}.
There is a pressing need for more robust characterization methods \cite{henryTinnitusEpidemiologicPerspective2020,henryMeasurementTinnitus2016,norenaPsychoacousticCharacterizationTinnitus2002,robertsResidualInhibitionFunctions2006}.

% INTRO PARAGRAPH 2/3
Tinnitus sounds are presumed to be complex and heterogeneous
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,henryTinnitusEpidemiologicPerspective2020,meikleTinnitusArchiveArchive2004},
although few specifics have been firmly established.
Therefore, we base our approach on \textit{reverse correlation} (RC),
an established behavioral method
\cite{ahumadaStimulusFeaturesSignal1971,gosselinSuperstitiousPerceptionsReveal2003,brimijoinInternalRepresentationVowel2013}
for probing internal representations of visual and auditory percepts, unconstrained by prior knowledge.
RC asks subjects to render subjective judgments over random stimuli, 
from which unbiased estimates of the percepts can be produced.
RC is closely related to Wiener theory \cite{wienerNonlinearProblemsRandom1966},
which has inspired ``white-noise'' approaches to system characterization in physiology
\cite{ringachReverseCorrelationNeurophysiology2004,marmarelisWhiteNoiseMethodSystem1978} and engineering \cite{ljungMeasureLackFit1978}.

% INTRO PARAGRAPH 3/3
Here, we validate RC as a method for characterizing all component
frequencies of non-tonal tinnitus sounds
by estimating the full psychoacoustic tinnitus spectrum (PTS).
To that end, we asked normal-hearing participants to complete an augmented RC experiment, comparing random stimuli to a target sound.
The estimated PTS were subsequently validated against the targets.
Our results demonstrate, for the first time, that tinnitus-like sounds with complex spectra can be accurately estimated using RC.

@@

\section{Materials and Methods}
% To use this template, type over sections of the template or cut and paste from another document. Highlight a section that you want to designate with a certain style, then select the appropriate name on the style menu. The style will adjust your fonts and line spacing. Do not change the font sizes or line spacing to squeeze more text into a limited number of pages. Use italics for emphasis; do not underline. 
% IEEE will do the final formatting of your paper. However, you should format your manuscript for submission to IEEE OJEMB for review and also for preparing for preprint that will be e-published several days after you upload the final version of an accepted manuscript.

% Software for the experiments and analysis was written in MATLAB.\footnote{Code is freely available at \protect\url{https://github.com/alec-hoyland/tinnitus-reconstruction/}}

@@

Stimuli were composed of $b=8$ Mel-spaced frequency bins,
which divide the frequency spectrum between $[100,~13,000]$
Hz into continuous segments of equal amplitude (\ie ``rectangular'' bins).
Reconstruction detail increases with $b$, but $b=8$ provides a good approximation to the chosen target sounds (see Fig. \ref{fig:spectra}).

For each stimulus, $[2,7]$ bins were randomly ``filled'' with power $0$ dB. ``Unfilled'' bins were assigned a power of $-100$ dB. 
Assigning random phase, the inverse Fourier transform of the spectrum yields a $500$ ms stimulus waveform.\footnote{Software for the experiments and analysis was written in MATLAB and is freely available at \protect\url{https://github.com/alec-hoyland/tinnitus-reconstruction/}}

@@

Two spectrally complex and complimentary target sounds (``buzzing'' and ``roaring'') were downloaded from the American Tinnitus Association \cite{MeasuringTinnitusAmerican2022} and truncated to $500$ ms in duration.

@@

10 normal-hearing subjects listened to the same target sound followed by a random stimulus for each trial (Fig. \ref{fig:experimentdiagram}).
The task is designed to mimic how a tinnitus patient might compare random stimuli to their tinnitus.

Subjects were instructed to respond ``yes'' to any stimuli that sounded similar to the target, and ``no'' otherwise.
Subjects completed $200$ trials for both target sounds within $10-20$ minutes.

@@

where $x \in \mathbb{R}^\mathrm{b}$ is the subject's internal representation of interest (\eg their tinnitus percept).
This model can be inverted, yielding the following formula:

@@

The experimental paradigm allows for direct validation of the reconstructions $\hat{x}_{buzzing}$ and $\hat{x}_{roaring}$.
We represent the spectra of the target sounds as vectors $s_{buzzing} \in \mathbb{R}^\mathrm{b}$ and $s_{roaring} \in \mathbb{R}^\mathrm{b}$ using the same frequency bins as in stimulus generation with power equal to the mean power of frequencies within that bin.
Pearson's $r$ between $s_{buzzing}$ and $s_{roaring}$ and their corresponding reconstructions represents reconstruction accuracy.
One-sample t-tests were performed on the mean Fisher-transformed Pearson's $r$ values across subjects to assess significant differences from zero.

\subsection{Synthetic Subjects}

To establish bounds on human performance,
additional experiments were run with two simulated subjects who give either \textit{ideal} or \textit{random} responses.
Each experiment ran for $n=200$ trials and was repeated $1000$ times.

@@

% NB NOTE: What is s??
for $i \in 1, ..., n$, where $Q(x, y)$ is the quantile function for $x \in [0, 1]$ of the similarity calculation $\Psi s$, and
$\Psi_i$ is the $i^\mathrm{th}$ column of $\Psi$.
The ideal subject has precise knowledge of every stimulus and responds mirroring Eq. \ref{eq:linreg}.

The \textit{random} subject responds with uniformly random probability.

@@

Figure \ref{fig:reconstructions} shows the distribution of Pearson's $r$ for human, ideal, and random subject responses.
Human accuracies were generally higher than random,
with some approaching ideal.
Mean random accuracy was $0$ in all conditions,
while mean human accuracy was significantly different from $0$ in all conditions: buzzing: $t(9) = 5.766$, $p < 0.001$; roaring: $t(9) = 5.76$, $p < 0.001$; combined: $t(19) = 7.542$, $p < 0.001$.
Greater variability in buzzing results may stem from the target spectrum containing significant high-frequency content, increasing the difficulty for subjects with minor high-frequency hearing loss.
Figure \ref{fig:spectra} plots the most accurate human reconstructions over the target sound spectra.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{gfx/reconstruction_violin_cropped.pdf}
    \caption{Human reconstruction accuracy is significantly above baseline,
    but is not optimal. Random, human, and ideal reconstruction accuracies
    are shown as violin plots with box plots overlaid.
    The median is a white dot, the ordinate of the blue points are the Pearson's $r$ values.}
    \label{fig:reconstructions}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{gfx/reconstruction_spectra_cropped.pdf}
    \caption{Reconstructions of the PTS capture many salient features of tinnitus spectra.
    The black trace indicates the target,
    while colored traces plot exemplar human reconstructions.
    The top row shows standardized power levels within each frequency bin.
    The bottom row maps the $8$-dimensional bin domain
    to a $11025$-dimensional frequency domain with the unbinned power spectral density of the targets shown in gray.}
    \label{fig:spectra}
\end{figure}

@@

Reconstruction accuracies observed here are well below the simulated ideal,
which may be attributed to noisy responses universally observed in applications of RC.
This may be mitigated by optimizing the experimental protocol, stimulus generation, and reconstruction method.
For example, recent approaches to improving RC reconstruction methods can boost efficiency, noise robustness and overall accuracy \cite{roopCompressiveSensingApproach2021,comptonStimulusWhiteningImproves2022a}.

Nevertheless, our results show that RC can accurately reconstruct the frequency spectrum of complex, non-tonal tinnitus-like sounds.
This indicates that RC may also be useful for developing a clinical assay to characterize a wider variety of tinnitus percepts than currently possible.
Additionally, subjects completed the required number of trials within 20 minutes,
indicating that this procedure could be conducted in a single clinical visit.
Future work will focus on validating this approach in patients suffering from tinnitus.
